### 2015-09-07 5_results_discussions/coselog-wabo/coselog-wabo.tex
no message

\subsection{Environmental Permit Application Process}
\label{subsec:coselog-wabo-process}